---
title: "Working title: Effect of symmetry on quantity discrimination"
author: "Leslie Ng"
date: "2024-01-08"
output:
  html_document:
   toc: yes
   toc_float: yes
   collapsed: true 
   smooth_scroll: true
   depth: 3 
   highlight: tango # different theme for the appearance of the code
   theme: flatly
   code_folding: hide
self_contained: yes
mode: selfcontained
editor_options:
  chunk_output_type: console
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../Output") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Introduction**

In this study we tested if training bees using symmetrical/asymmetrical stimuli had an effect on their quantity discrimination. This document provides details on data processing, statistical analysis and figures of the original manuscript submitted for peer review.

The following code is written in the R programming language.

# **Install required packages**

```{r eval=FALSE}
install.packages("lme4") # linear mixed effect models
install.packages("ggplot2") # plot graphs
install.packages("tidyverse") # QOL features
install.packages("wesanderson") # color palette
install.packages("dplyr")  # data manipulation
install.packages("BFpack") # Bayes Factor for logistic regression
install.packages("ggThemeAssist") # Add-in to customize ggplot2 themes
install.packages("performance") # For assumption checking
install.packages("DHARMa") # For assumption checking]
install.packages("patchwork") # For combining figures together
install.packages("MuMIn")
```

# **Load packages**

```{r, message=FALSE, warning=FALSE}
library(lme4) # linear mixed effect models
library(ggplot2) # plot graphs
library(tidyverse) # QOL features
library(wesanderson) # color palette
library(dplyr)  # data manipulation
library(BFpack) # Bayes Factor for logistic regression
library(ggThemeAssist) # you may need to restart you session for this add-in to appear
library(performance) # For assumption checking
library(DHARMa) # For assumption checking
library(patchwork)
library(emmeans)
library(MuMIn)
```

# **Data importing**

```{r, message=FALSE, results="hide"}

# Data importing -----------------------------------------------------------------------

test <- read.csv("../Data/sym_test.csv", header=TRUE)
training <- read.csv("../Data/sym_training.csv", header=TRUE)
summary(test) #view summary
summary(training)


# Data wrangling ------------------------------------------------------------------------

test$BEEID <- as.factor(test$BEEID) # treat as a categorical factor
test$TESTCONDITION <- as.factor(test$TESTCONDITION)
test$TRAININGCONDITION <- as.factor(test$TRAININGCONDITION)

training$Bee_ID <- as.factor(training$Bee_ID) # treat as a categorical factor
training$Group <- as.factor(training$Group) # treat as a categorical factor
training$Group_letter <- as.factor(training$Group_letter)

trainingsym <-subset(training, Group_letter
 == "S")
trainingasym <-subset(training, Group_letter
 == "A")


symmetry <-subset(test, TRAININGCONDITION == "SYM") #subset data
asymmetry <-subset(test, TRAININGCONDITION == "ASYM")

symsym <-subset(symmetry, TESTCONDITION == "SYM") #subset data
mean(symsym$CHOICE)

asymsym <-subset(asymmetry, TESTCONDITION == "SYM")
mean(asymsym$CHOICE)

symasym <-subset(symmetry, TESTCONDITION == "ASYM") #subset data
mean(symasym$CHOICE)

asymasym <-subset(asymmetry, TESTCONDITION == "ASYM")
mean(asymasym$CHOICE)

symtrans <-subset(symmetry, TESTCONDITION == "TRANSF") #subset data
mean(symtrans$CHOICE)

asymtrans <-subset(asymmetry, TESTCONDITION == "TRANSF")
mean(asymtrans$CHOICE)

symconflict <-subset(symmetry, TESTCONDITION == "CONFL") #subset data
mean(symconflict$CHOICE)

asymconflict <-subset(asymmetry, TESTCONDITION == "CONFL")
mean(asymconflict$CHOICE)

symconflict2 <-subset(symmetry, TESTCONDITION == "CONFLLARGE") #subset data
mean(symconflict2$CHOICE)

asymconflict2 <-subset(asymmetry, TESTCONDITION == "CONFLLARGE")
mean(asymconflict2$CHOICE)



STRAINING <-subset(training, Group_letter == "S") #subset data
ATRAINING <-subset(training, Group_letter == "A")
symmetrytest <-subset(test, TESTCONDITION == "SYM") #subset data
asymmetrytest <-subset(test, TESTCONDITION == "ASYM") #subset data
transtest <-subset(test, TESTCONDITION == "TRANSF") #subset data

```

# **Logistic regressions**

### *Training data*

```{r message=FALSE, error=FALSE}

##Full model including all variables and both groups##

bees_training_model_FULL <- glmer(Choice ~ Trial + Group + Low_number + High_number + Shape_number + (1|Bee_ID), data = training, family = binomial)
summary (bees_training_model_FULL)

options(na.action=na.fail) #dredge function to find ideal model and variables to include
Bee_train_model_dredge<-dredge(bees_training_model_FULL)
Bee_train_model_dredge
options(na.action = "na.omit") ##The issue with using options(na.action = "na.fail") is that it changes the global settings of R. If you have a large script, changing the global settings will potentially impact on other sections of your code where you implicitly rely on Rs default settings. There are two ways to avoid this: After using dredge change the settings back via options(na.action = "na.omit").

trainingglm <- glmer(Choice~ Trial*Group_letter + (1|Bee_ID), family = binomial, data = training)
summary(trainingglm)

emmeans(trainingglm, ~ Group_letter | Trial, at = list(Trial = c(1, 50, 100)), type = "response")

##Group comparison##


##Training analysis symmetry and asymmetry groups seperately##
Symmetry_group_training <- subset(training, Group == 1) #subset symmetry group
summary(Symmetry_group_training)
Asymmetry_group_training <- subset(training, Group == 2) #subset asymmetry group
summary (Asymmetry_group_training)

Sym_bees_training_model_FULL <- glmer(Choice ~ Trial + Group + Low_number + High_number + Shape_number + (1|Bee_ID), data = Symmetry_group_training, family = binomial)
summary (Sym_bees_training_model_FULL)
options(na.action=na.fail) #dredge function to find ideal model and variables to include
Bee_train_model_sym_dredge<-dredge(Sym_bees_training_model_FULL)
Bee_train_model_sym_dredge
options(na.action = "na.omit") 
symmetry_bees_training_model <- glmer(Choice ~ Trial + High_number + (1|Bee_ID), data = Symmetry_group_training, family = binomial)
summary (symmetry_bees_training_model)

Asym_bees_training_model_FULL <- glmer(Choice ~ Trial + Group + Low_number + High_number + Shape_number + (1|Bee_ID), data = Asymmetry_group_training, family = binomial)
summary (Asym_bees_training_model_FULL)
options(na.action=na.fail) #dredge function to find ideal model and variables to include
Bee_train_model_asym_dredge<-dredge(Asym_bees_training_model_FULL)
Bee_train_model_asym_dredge
options(na.action = "na.omit") 
Asymmetry_bees_training_model <- glmer(Choice ~ Trial + (1|Bee_ID), data = Asymmetry_group_training, family = binomial)
summary (Asymmetry_bees_training_model)

```



### *Test data*
The first part of the model name indicates the training condition, and the second part indicates the specific test e.g. symasymlm refers to the model on data for the asymmetrical test involving bees trained with symmetrical stimuli 
```{r message=FALSE, error=FALSE}
symsymlm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = symsym)
summary(symsymlm)

asymsymlm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = asymsym) 
summary(asymsymlm)

symasymlm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = symasym)
summary(symasymlm)

asymasymlm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = asymasym) 
summary(asymasymlm)

symtranslm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = symtrans)
summary(symtranslm)

asymtranslm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = asymtrans) 
summary(asymtranslm)

symconfllm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = symconflict)
summary(symconfllm)

asymconfllm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = asymconflict) 
summary(asymconfllm)

symconfllm2 <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = symconflict2)
summary(symconfllm2)

asymconfllm2 <-glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = asymconflict2) 
summary(asymconfllm2)

symtestcompare<- glmer(CHOICE~TRAININGCONDITION+(1|BEEID), family = binomial, data = symmetrytest) 
summary(symtestcompare)

asymtestcompare<- glmer(CHOICE~ TRAININGCONDITION+(1|BEEID), family = binomial, data = asymmetrytest) 
summary(asymtestcompare)

transfertestcompare<- glmer(CHOICE~ TRAININGCONDITION+(1|BEEID), family = binomial, data = transtest) 
summary(transfertestcompare)

levels(test$TESTCONDITION)
levels(test$TRAININGCONDITION)

test$TESTCONDITION <- relevel(test$TESTCONDITION, ref = "SYM")
test$TRAININGCONDITION <- relevel(test$TRAININGCONDITION, ref = "SYM")

testtypeglm<- glmer(CHOICE~TESTCONDITION*TRAININGCONDITION+(1|BEEID), family = binomial, data = test) 
summary(testtypeglm)
testtypeglm<- glmer(CHOICE~TESTCONDITION+TRAININGCONDITION+(1|BEEID), family = binomial, data = test) 
summary(testtypeglm)
emmeans_test <- emmeans(testtypeglm, ~ TESTCONDITION)
pairs(emmeans_test)

```

### *Confidence intervals*

```{r results='hide', error=FALSE, message=FALSE, warning=FALSE}

confint(trainingglm)
trainingCNUM2confintlow<-(exp(0.26402783)/(1 + exp(0.26402783))) #reverse link function
trainingCNUM2confinthigh<-(exp(0.47046562)/(1 + exp(0.47046562)))
trainingSYMSconfintlow<-(exp(-0.27500042)/(1 + exp(-0.27500042))) #reverse link function
trainingSYMSconfinthigh<-(exp(0.12237367)/(1 + exp(0.12237367)))
trainingINTERconfintlow<-(exp(-0.37732274)/(1 + exp(-0.37732274))) #reverse link function
trainingINTERconfinthigh<-(exp(-0.09007661)/(1 + exp(-0.09007661)))

confint(symsymlm) #confidence interval
symsymlmconfintlow<-(exp(0.1684812)/(1 + exp(0.1684812))) #reverse link function
symsymlmconfinthigh<-(exp(0.8137388)/(1 + exp(0.8137388)))

confint(asymsymlm)
asymsymlmconfintlow <-(exp(0.2992044)/(1 + exp(0.2992044)))
asymsymlmconfinthigh<-(exp(1.166141 )/(1 + exp(1.166141)))

confint(symasymlm)
symasymlmconfintlow <-(exp(0.1243281)/(1 + exp(0.1243281)))
symasymlmconfinthigh<-(exp(0.9110778 )/(1 + exp(0.9110778)))

confint(asymasymlm)
asymasymlmconfintlow <-(exp(0.536397)/(1 + exp(0.536397)))
asymasymlmconfinthigh<-(exp(1.456451 )/(1 + exp(1.456451)))

confint(symtranslm)
symtranslmconfintlow <-(exp(0.3327167)/(1 + exp(0.3327167)))
symtranslmconfinthigh<-(exp(1.188838 )/(1 + exp(1.188838)))

confint(asymtranslm)
asymtranslmconfintlow <-(exp(0.5579531)/(1 + exp(0.5579531)))
asymtranslmconfinthigh<-(exp(1.3551496 )/(1 + exp(1.3551496)))

confint(symconfllm)
symconfllmconfintlow <-(exp(-0.3037767)/(1 + exp(-0.3037767)))
symconfllmconfinthigh<-(exp(0.3553857)/(1 + exp(0.3553857)))

confint(asymconfllm)
asymconfllmconfintlow <-(exp(-0.1840665)/(1 + exp(-0.1840665)))
asymconfllmconfinthigh<-(exp(0.747081)/(1 + exp(0.747081)))

confint(symconfllm2)
symconfllm2confintlow <-(exp(-0.6812548)/(1 + exp(-0.6812548)))
symconfllm2confinthigh<-(exp(1.239418)/(1 + exp(1.239418)))

confint(asymconfllm2)
asymconfllm2confintlow <-(exp(-0.5036269)/(1 + exp(-0.5036269)))
asymconfllm2confinthigh<<-(exp(0.3303297)/(1 + exp(0.3303297)))

```

# **Generate figures**

### *Create dataframe for figures*

```{r message=FALSE, results='hide', error=FALSE}
symdframe <- data.frame(condition = rep(c("Symmetry","Asymmetry","Transfer","Conflict small","Conflict large")), # Create dataframe
                legend = c("S","A","T","C-S","C-L"), 
                proportion = c(0.619,0.619,0.669,0.506,0.54),
                confintlow = c(symsymlmconfintlow,symasymlmconfintlow,symtranslmconfintlow,symconfllmconfintlow,symconfllm2confintlow),
                confinthigh = c(symsymlmconfinthigh,symasymlmconfinthigh,symtranslmconfinthigh,symconfllmconfinthigh,symconfllm2confinthigh))



symmeanpoints <- symmetry %>%   # Calculate mean of each block for each bee
  group_by(BEEID, TESTCONDITION) %>% 
  summarize(prop = mean(CHOICE)) %>%
  mutate(prop = ifelse(prop == 1, 0.95, prop)) %>%
  mutate(prop = ifelse(prop == 0, 0.05, prop))
  colnames(symmeanpoints)[2] ="condition" # rename column to match dframe
  levels(symmeanpoints$condition) <- tolower(levels(symmeanpoints$condition)) # rename condition names to match 
  levels(symmeanpoints$condition) <- str_to_title(levels(symmeanpoints$condition)) # recapitalize condition names
  levels(symmeanpoints$condition)
  levels(symmeanpoints$condition) <- c('Asymmetry', 'Conflict small', 'Conflict large', 'Symmetry', 'Transfer')
  symmeanpoints$condition <- factor(symmeanpoints$condition, levels=c('Symmetry', 'Asymmetry', 'Transfer', 'Conflict small', 'Conflict large'))
  levels(symmeanpoints$condition)
  symmeanpoints <- subset(symmeanpoints, condition == "Symmetry" | condition == "Asymmetry" | condition == "Transfer" | condition == "Conflict small" | condition == "Conflict large")
symdframe$condition <- factor(symdframe$condition, levels = c('Symmetry', 'Asymmetry', 'Transfer', 'Conflict small', 'Conflict large'))

                           
  
asymdframe <- data.frame(condition = rep(c("Symmetry","Asymmetry","Transfer","Conflict small","Conflict large")), # Create dataframe
                legend = c("S","A","T","C-S","C-L"), 
                proportion = c(0.663,0.713,0.713,0.563,0.48),
                confintlow = c(asymsymlmconfintlow,asymasymlmconfintlow,asymtranslmconfintlow,asymconfllmconfintlow,asymconfllm2confintlow),
                confinthigh = c(asymsymlmconfinthigh,asymasymlmconfinthigh,asymtranslmconfinthigh,asymconfllmconfinthigh,asymconfllm2confinthigh))

asymmeanpoints <- asymmetry %>%   # Calculate mean of each block for each bee
  group_by(BEEID, TESTCONDITION) %>% 
  summarize(prop = mean(CHOICE)) %>%
  mutate(prop = ifelse(prop == 1, 0.95, prop)) %>%
  mutate(prop = ifelse(prop == 0, 0.05, prop))
  colnames(asymmeanpoints)[2] ="condition" # rename column to match dframe
  levels(asymmeanpoints$condition) <- tolower(levels(asymmeanpoints$condition)) # rename condition names to match 
  levels(asymmeanpoints$condition) <- str_to_title(levels(asymmeanpoints$condition)) # recapitalize condition names
  levels(asymmeanpoints$condition) <- c('Asymmetry', 'Conflict small', 'Conflict large', 'Symmetry', 'Transfer')
  asymmeanpoints$condition <- factor(asymmeanpoints$condition, levels=c('Symmetry', 'Asymmetry', 'Transfer', 'Conflict small', 'Conflict large'))
  levels(asymmeanpoints$condition)
  asymmeanpoints <- subset(asymmeanpoints, condition == "Symmetry" | condition == "Asymmetry" | condition == "Transfer" | condition == "Conflict small" | condition == "Conflict large")
asymdframe$condition <- factor(asymdframe$condition, levels = c('Symmetry', 'Asymmetry', 'Transfer', 'Conflict small', 'Conflict large'))



newdata <- expand.grid(
  Trial = seq(min(training$Trial), max(training$Trial), length.out = 100),
  Group_letter = levels(training$Group_letter)
)
training$Group_letter <- as.factor(training$Group_letter)


pred <- predict(trainingglm, newdata = newdata, type = "link", se.fit = TRUE, re.form = NA)

newdata$fit <- plogis(pred$fit)
newdata$lower <- plogis(pred$fit - 1.96 * pred$se.fit)
newdata$upper <- plogis(pred$fit + 1.96 * pred$se.fit)



```

# **Generate figure**

```{r}


symgraph <- ggplot(symdframe, aes(x = condition, y = proportion, fill = condition)) +
  labs(x = "Test condition", y = "Mean proportion of choices", fill = "Condition") +
  geom_bar(stat = "identity", position = "dodge", width = .7) + 
  geom_point(data = symmeanpoints, aes(x = condition, y = prop, colour = BEEID), 
             position = position_jitter(width = 0.15, height = 0.01), size = 2) +
    scale_fill_manual(values = wes_palette("Moonrise3", n = 5)) +
  geom_hline(yintercept = 0.5, linetype = "dashed", colour = "black") +
  geom_errorbar(aes(ymin = confintlow, ymax = confinthigh), width = .2) +
  geom_text(aes(x = 'Symmetry', y = 0.99, label = "**"), 
          inherit.aes = FALSE, size = 10, colour = "black") +
  geom_text(aes(x = 'Asymmetry', y = 0.99, label = "**"), inherit.aes = FALSE, size = 10, colour = "black") +
  geom_text(aes(x = 'Transfer', y = 0.99, label = "***"), inherit.aes = FALSE, size = 10, colour = "black") +
  scale_y_continuous(limits = c(0, 1)) +
    scale_x_discrete(labels = c("Sym", "Asym", "Tran", "Con4","Con10")) +  
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        aspect.ratio = 1 / 1,
        legend.position = "none",
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 13),
        plot.margin = margin(0.3,0.3,0.3,0.6, "cm"),
        axis.title.y = element_text(vjust = 3),
        axis.title.x = element_text(margin = margin(t = 10)))
print(symgraph)



asymgraph <- ggplot(asymdframe, aes(x = condition, y = proportion, fill = condition))+
  labs (x= "Test condition", y = "Mean proportion of choices", fill="Condition") +
  geom_bar(stat = "identity", position = "dodge", width = .7) + 
  geom_point(data = asymmeanpoints, aes(x = condition, y = prop, colour = BEEID), position = position_jitter(width=0.15, height=0.01), size =2) +
  scale_fill_manual(values = wes_palette("Moonrise3", n = 5)) +
  geom_hline(yintercept = 0.5, linetype = "dashed", colour = "black") +
  geom_errorbar(aes(ymin = confintlow, ymax = confinthigh), width=.2) +
  geom_text(aes(x = 'Symmetry', y = 0.99, label = "***"), 
          inherit.aes = FALSE, size = 10, colour = "black") +
  geom_text(aes(x = 'Asymmetry', y = 0.99, label = "***"), inherit.aes = FALSE, size = 10, colour = "black") +
  geom_text(aes(x = 'Transfer', y = 0.99, label = "***"), inherit.aes = FALSE, size = 10, colour = "black") +
  scale_y_continuous(limits = c(0,1)) +
  scale_x_discrete(labels = c("Sym", "Asym", "Tran", "Con4","Con10")) +  # Change x-axis labels +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        aspect.ratio = 1/1,
        legend.position = "none",
                axis.title = element_text(size = 15),
        axis.text = element_text(size = 13),
        axis.title.y = element_blank(),
        plot.margin = margin(0.3,0.3,0.3,0.3, "cm"),
        axis.title.x = element_text(margin = margin(t = 10)))

print(asymgraph)

trainingfig <- ggplot(newdata, aes(x = Trial, y = fit, color = Group_letter, group = Group_letter)) +
  geom_line(size = 1.2, aes(linetype = Group_letter)) +  
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) + 
  labs(x = "Trial number", y = "Mean predicted proportion of correct choices") +
  scale_color_manual(values = c("skyblue", "salmon")) +  # Custom colors for each condition
  scale_y_continuous(breaks = seq(0, 1, by = 0.1)) +     # Set y-axis ticks every 0.1
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "none",
    panel.background = element_blank(),  # Removes background
    panel.grid = element_blank(),        # Removes grid lines
    axis.line = element_line(),          # Adds axis lines
    axis.ticks = element_line(),
     aspect.ratio = 1/1,
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 15),# Ensures ticks are displayed
  )

print(trainingfig)


```

### *Saving combined figure*

```{r}

figx <- symgraph + asymgraph
figx + plot_annotation(tag_levels = "A") & 
  theme(plot.tag.position = c(-0.1, 1),plot.tag = element_text(size = 16, face = "bold"))
ggsave("fig4.pdf", device = pdf, height= 4.1, width=8.5, dpi = 300, path = "../Output")
ggsave("fig4.png", device = png, height= 4.1, width=8.5, dpi = 300, path = "../Output")

trainingfig
ggsave("training.png", device = png, height= 6.2, width=6.2, dpi = 300, path = "../Output")
```

# **Check Model fit for above models**

```{r warning=FALSE, error = FALSE, message=FALSE}

models <- list(
  symsym = symsymlm,
  asymsym = asymsymlm,
  symasym = symasymlm,
  asymasym = asymasymlm,
  symtrans = symtranslm,
  asymtrans = asymtranslm,
  symconfl = symconfllm,
  asymconfl = asymconfllm,
  symtestcompare = symtestcompare,
  asymtestcompare = asymtestcompare,
  transfertestcompare = transfertestcompare,
  training = trainingglm,
  symmetrytraining = symmetry_bees_training_model,
  asymmetrytraining = Asymmetry_bees_training_model,
  testtype = testtypeglm
)

# Check for overdispersion
for (model_name in names(models)) {
  cat(paste0("Overdispersion Check for ", model_name, "\n"))
  print(check_overdispersion(models[[model_name]]))
  cat("\n")
}

# Compute performance metrics for each model
performance_results <- lapply(models, model_performance)

# Capture and print markdown output for each model
for (model_name in names(performance_results)) {
  cat(paste0("### Performance Metrics for ", model_name, "\n\n"))
  
  # Capture the output in markdown format
  output <- capture.output(display(performance_results[[model_name]], format = "markdown", digits = 2, caption = NULL))
  
  # Print the captured output to the console
  cat(paste(output, collapse = "\n"))
  
  cat("\n\n")
}

# Loop for simulating residuals
for (model_name in names(models)) {
  cat(paste0("Simulated Residuals and Plot for ", model_name, "\n"))
  
  # Simulate residuals
  simulated_res <- simulateResiduals(fittedModel = models[[model_name]])
  
  # Plot simulated residuals
  plot(simulated_res)
  
  cat("\n\n")
}

```
